NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-06-26 paper/itemizedReviewResponse.txt

Q1. The usage scenarios also look interesting, especially the Scenario 3. Does the result to some extent suggest that the attention mechanism is not a very efficient component in this task? I expect the authors dig it further.

A: The domain experts indeed find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From the result from Scenario 3, we can extrapolate that the ultimate decision relies on the semantics (i.e., encodings and composed encodings in classifier). The encoder and classifier layer can swiftly adjust weights for word/composed semantics to correct prediction. But the attention layer, which only works on how individual words are aligned, affects the prediction less significantly. Recent works also support this observation in word embeddings, e.g., ELMo by Salant et al., where better word embeddings can substantially benefit a model. However, it by no means implies attention is not useful, as it serves as a way to compose word semantics although it does not directly output that. We revised section 6.3 to include this comment on the interpretation of the visualization result.

Q2. In Section 2, the neural net in NLP and attention mechanisms are introduced. But the descriptions are unsatisfactory. There is no formal definition or description of the model and the attention mechanism. It's still unclear how the attention mechanism is applied in the model. I would suggest the authors include a concise and formal introduction in the background part and include some formula.

A: The exact computation of attention can vary from model to model despite the shared overall concept, i.e., the alignment between words. As a result, when we discuss the general concept of attention, we cannot fix on a single mathematical formulation. The exact formulation for the model we investigated is discussed in Appendix A.

@@

Q3, In Section 4, the paper indicates that the system is developed under a long-term collaboration with NLP experts, but there is no description describing the details of the collaboration (the number of experts, the length and the form of the collaboration, the frequency of the meetings). It's not until Section 7 did I learn that there are 2 NLP experts in collaboration. The task analysis can also be improved by including more detailed justification of the three tasks.

A: We had added additional details of the collaboration at the beginning of section 4, including the number of NLP collaborators, frequency, duration of the meeting period for the proposed work. And we also revised the motivation for these tasks and pointed out their connection with the scenarios discussed in section 6.

@@

Q7, It's better if the authors present the relation between Section 5.3 and 5.4 in a clearer way. It's confusing to read how one can perturb a prediction in Section 5.3 before he/she read the algorithm in Section 5.4. I suggest put the algorithm along with the perturbation-driven paradigm at the beginning of Section 5.

A: We thank the reviewer for catching the ordering problem and providing a solution. To address the issue, we added references to the optimization discussed in section 5.4 at the beginning of section 5 as well as in section 5.3. We didn't move all the optimization discussion at the beginning of section 5, since we believe presenting details of the all three perturbations (input, attention, prediction) without any discussion about the interface can also be very confusing. And the separation between the operation and UI component may also make the description harder to read (require the reader to refer back to the computation discussion when reading about the user interface). We believe the revised version provides a decent tradeoff.

Q8, Except for the presentation issue, the contribution of this paper is somewhat limited to only NLI models with attention mechanisms. It's not sure how the visualization can be easily extended for other NLP tasks. For example, I am not sure how the proposed visualization can easily adapt to machine translation, considering that there are no neutral/contradiction/entailment labels.

A: The composability of the system is the key to extend its capability. The proposed API allows the user to include only the view that is necessary for the given task. For a new task (i.e., machine translation), we do not need to include the prediction view for the NLI task. We can reuse the attention views for representing the alignment between the original sentence and the translated sentence and introduce new views for visualizing other information that may not present in the NLI model. The main idea is to allow the user to combine the view/functionality and add new view if necessary to suit their analysis goals.

@@

Q9, In addition, the perturbation-driven scheme is somewhat related to the gradient-based analysis (both are the local analysis of the model on a specific prediction). It is better if the authors discuss the two and differentiate them.

A: Gradient-based sensitivity analysis can often be carried out for a domain that is differentiable (i.e., image pixel values). In the proposed work, the challenge of using the gradient-based method in NLP inspired us to utilize perturbation. This point is partly discussed in the introduction where the difference between NLP and CV is highlighted (image has a continuous domain whereas words are discrete).

@@

Q10, The evaluation is acceptable but a bit weak. There is no description of what questions have the authors asked during the evaluation sessions. It would be better to have formal interviews and ask the participants to perform specific tasks and observe how they performed.

A: We didn't design a specific list of tasks and the corresponding answers mostly due to the open-ended analysis goal. If the user didn't follow the particular exploration path the designer of the question goes, he/she might find some result that not directly related to the "correct" answer. However, that does not mean the analysis is wrong, they just approach the data from a different perspective. However, we do believe extensive evaluation is extremely important for improving the usability of the tool. We plan to recruit more domain experts to test and evaluate the system in the future.

@@

Q38, It was unclear to me if the idea of perturbing a model is novel, or an extension/visualization of previously introduced work in this area applied to NLP NNs. If it's the former (as it seems to be), then I think this needs to be emphasized in the introduction.

A: The perturbation of nouns and verbs on itself is not novel from NLP perspective. And truly solving the challenge of producing grammatically correct and meaningful perturbation/paraphrasing is still an open research area. In our case, we believe the novelty comes from the integration of the model with the three-stage perturbation scheme in an explorable visual analytic system. We have revised the introduction to highlight the core contribution.

Q39, I appreciated the task analysis in section 4. However, it seems a bit too brief, and the tasks don't seem to directly follow from the description of the authors experience with the NLP researchers. The scenarios described in Section 6 provide much more detail about these tasks, and Section 5 clarifies them as well, but a bit more detail here would be useful. For example, it's not clear who the NLP researchers are, how many of them were interviewed, what type of NLP projects they were working on, etc.

A: We moved the description of the collaboration into the beginning of section 4, in which we described the number of NLP experts, frequency, and the duration of the meetings for this specific project. We also added a reference to application scenarios that correspond to the tasks to help instantiate the more abstract concepts.

@@

Q40, Work by Ma et al. ("Visualizing dynamic brain networks using an animated dual-representation" from EuroVis 2015) is work related to NLIZE having both graph and matrix views that complement each other and provide the same information from different perspectives, as described in Section 5.2.

A: We've cited this work in Section 5.2 when discussing the due visual representations.

Q41, Placing the grammar dependency tree over the matrix / bipartite graph is interesting (described in Section 5 and Section 6.4). I wonder if it would make sense to use other types of annotations here, such as those created from event extraction tools or information extraction tools that would highlight other structures (besides grammar/POS) that could be useful in helping understanding the inference model. Work by Forbes et al. supports the use of semantic hypergraphs to provide richer textual annotations alongside text sequences ("Text annotation graphs: Annotating complex natural language phenomena" from LREC 2018).

A: After discussing with domain experts, we believe the reviewer's comment would be an exciting direction for future works. Coming up with efficient visual encoding for these types of information that go beyond grammar tree is a challenge on its own. For example, the extract structure can be a graph (instead of a tree), which likely requires a complete redesign of the existing visualization.

@@

Q42, In Sections 6.1 and 6.2, the application scenario gives an example of having a user perturbing the model by changing the input words (switching "pile" with "heap"). This updates the system and changes the inference output. Are the successful results of such an interaction remembered by the system? That is, could NLIZE be used to tweak/update the model so that, say, the ambiguity between pile and heap is automatically and more accurately resolved in future queries?

A: In the current system, we do not record the user interaction/correction result. The MIRA optimization update may correct the model for the given example (to some extend overfit to the example). However, we don't think there is a guarantee that the fix can be readily generalized to other cases. Also, the ambiguity between pile and heap may also originate from the word embedding (the numerical representation for words). The proposed system suggests to NLP experts that the uncertainty between pile and heap is likely the cause of the incorrect prediction. The experts can choose to dive further and examine the exact cause.

### 2018-06-26 paper/intro.tex
minor fix

% (What is the high level motivation)
% What is the obstacle for effectively apply machine learning model
%Based on some stunning recent advances, neural networks based machine learning approaches have gained unprecedented popularity, and have been adopted in a wide variety of applications.
As demonstrated by many recent successes, neural network-based machine learning approaches have garnered increasing popularity and have been adopted in a wide variety of applications.
%
However, researchers as well as practitioners often need to overcome many obstacles during the training, debugging, and tuning processes to realize the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical for both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and propose hypotheses for the cause of the failure is key to identifying the limitations of and improving upon existing models.
However, providing meaningful answers to these questions is challenging and has been described as impossible by many.

>>>>

% (What is the high level motivation)
% What is the obstacle for effectively apply machine learning model
%Based on some stunning recent advances, neural networks based machine learning approaches have gained unprecedented popularity, and have been adopted in a wide variety of applications.
As demonstrated by many recent successes, neural-network-based machine learning approaches have garnered increasing popularity and have been adopted in a wide variety of applications.
%
However, researchers as well as practitioners often need to overcome many obstacles during the training, debugging, and tuning processes to realize the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical for both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and propose hypotheses for the cause of the failure is key to identifying the limitations of and improving upon existing models.
However, providing meaningful answers to these questions is challenging and has been described as impossible by many.
